\documentclass[../main.tex]{subfiles}

\begin{document}
\section{Conclusion}
\label{sec:conclusion}
In this paper, we analyze the effect of an ageing population on the natural rate of interest in Denmark. We calibrate an overlapping generations model of an open economy with a realistic pension system and detailed demographic data. We find that the natural rate of interest has fallen by 1.8 percentage points in Denmark from 1985 to 2020. The demographic transition accounts for approximately half of the decline, while the other half is mainly due to sluggish global TFP growth. Higher life expectancy is the most important demographic factor accounting for approximately 0.7 percentage points of the decline in the NRI and low fertility by another 0.2 percentage points.

This downward pressure on the NRI has contributed to Denmark's existing low interest rate environment. According to our results, the demographic transition is projected to put further downward pressure on the NRI, increasing the probability of low interest rate environments in the future and strengthening the need for unconventional monetary policy and pension system reforms.

We have primarily focused on the pension system and analyzed the impact of increasing the retirement age and decreasing pension benefits suggested by the Danish government. Both suggestions will contribute to ensuring fiscal sustainability, but they have opposite implications on the real interest rate. Increasing the retirement age will alleviate the downward pressure on the NRI, but decreasing the replacement rate will contribute to an even lower NRI. 
\end{document}